---
title: 'A multiverse analysis of the associations between internet use and well-being'
shorttitle: Internet and well-being
leftheader: Internet and well-being
author: 
  - name: Matti Vuorre
    affiliation: "1,2"
    corresponding: yes
    address: Tilburg University
    email: mjvuorre@uvt.nl
  - name: Andrew K. Przybylski
    affiliation: "2"
    email: andy.przybylski@oii.ox.ac.uk
affiliation:
  - id: 1
    institution: Tilburg School of Social and Behavioral Sciences, Tilburg University
  - id: 2
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  \textbf{This manuscript is not yet peer-reviewed. The preprint is available at \url{https://psyarxiv.com/jp5nd/}.}
  We used the proprietary Gallup World Poll dataset, which is available to subscribing institutions through the Gallup website. All our code and a simulated GWP dataset is available at \url{https://doi.org/10.5281/zenodo.7774923}. We did not pre-specify the sample sizes, but instead used all the available data in the Gallup database (with the exclusions detailed below). This study was not pre-registered.
keywords: well-being, internet technology, technology effects
wordcount: "4,217"
bibliography: references.bib
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: doc
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
    keep_tex: false
---

```{r}
#| label: setup
#| include: false

library(papaja)
library(scales)
library(knitr)
library(lme4)
library(ggh4x)
library(emmeans)
library(ggside)
library(ggbeeswarm)
library(broom.mixed)
library(patchwork)
library(ggdist)
library(corrr)
library(tidyverse)
opts_chunk$set(
  include = FALSE,
  cache = TRUE
)

# Environment options
DATA_PATH_GWP <- Sys.getenv(
  "DATA_PATH_GWP"
)
DATA_PATH_GWP_CLEAN <- "data-raw/gwp.rds"
DATA_PATH_SYN <- Sys.getenv(
  "DATA_PATH_SYN"
)
MAX_CORES <- as.numeric(
  Sys.getenv(
    "MAX_CORES", 
    unset = parallel::detectCores(logical = FALSE)
  )
)

# Figure theme
theme_set(
  theme_linedraw(base_size = 9.6) +
    theme(
      strip.background = element_rect(fill = "gray90", color = "gray90"),
      strip.text = element_text(color = "black", hjust = 0),
      panel.grid = element_blank()
    )
)

# Function to clean up variable names for plots and tables
sanitize_labels <- function(x, pad7 = TRUE) {
  # Make internet predictors look ok
  x <- x %>% 
    str_replace("_", " ") %>% 
    str_to_sentence() %>% 
    str_replace("Gwb", "GWBI")
  if (pad7) x <- str_replace(x, "7", " 7 ")
  x
}

# Function to print large numbers
number2 <- function(x, ...) number(x, ..., big.mark = ',')
```

```{r}
#| label: data-clean

# Wrangle GWP .sav file
if (file.exists(DATA_PATH_GWP) & !file.exists(DATA_PATH_GWP_CLEAN)) {

  # List of indices
  ix <- list(
    life_satisfaction = c("WP16"),
    negative_experiences = c("WP68", "WP69", "WP70", "WP71", "WP74"),
    positive_experiences = c("WP60", "WP61", "WP63", "WP65", "WP67"),
    social_life = c("WP27", "WP10248"), 
    foodshelter = c("WP40", "WP43"), 
    gwb_purpose = c("WP14442", "WP14443"),
    gwb_community = c("WP14450", "WP14451"),
    gwp_social = c("WP14444", "WP14445"),
    gwb_physical = c("WP14448", "WP14449")
  )
  
  library(haven)
  library(labelled)
  # Load relevant variables
  dat <- read_spss(
    DATA_PATH, 
    col_select = c(
      WGT,
      WP5, 
      YEAR_CALENDAR,
      WP1220,
      WP1219, 
      WP3117,
      WP23, 
      WP1223, 
      EMP_2010,
      INCOME_2,
      WP2319,
      WP16056,
      WP39,
      WP19544,
      WP15862,
      all_of(
        unlist(ix)
      )
    )
  )
  
  # Replace values with labels DK and Refused to proper NAs
  val_labels(dat)  # These are the current value labels
  to_na <- function(x) {
    na_values(x) <- val_labels(x)[c("(DK)", "(Refused)")]
    x
  }
  dat <- dat %>% 
    mutate(across(c(everything(), -WP1219, -WP1220, -EMP_2010, -WP3117), to_na))
  na_values(dat$WP3117) <- val_labels(dat$WP3117)[c("(DK)", "(RF)")]
  dat <- user_na_to_na(dat)
  val_labels(dat)
  str(dat)
  
  dat <- dat %>% 
    mutate(
      country = fct_drop(as_factor(WP5)),
      year = as.integer(YEAR_CALENDAR),
      age = as.integer(ceiling(WP1220)),
      sex = factor(as_factor(WP1219), levels = c("Female", "Male")),
      weight = WGT,
      healthproblems = factor(as_factor(WP23), levels = c("No", "Yes")),
      relationship = fct_collapse(
        as_factor(dat$WP1223),
        Single = c(
          "Single/Never been married", 
          "Separated", 
          "Divorced", 
          "Widowed"
        ),
        Relationship = c("Married", "Domestic partner")
      ),
      income = as.numeric(INCOME_2),
      education = as.numeric(WP3117) - 1,
      work = EMP_2010,
      internet_access = factor(as_factor(WP16056), levels = c("No", "Yes")),
      internet_past7days = factor(as_factor(WP15862), levels = c("No", "Yes")),
      internet_home = factor(as_factor(WP39), levels = c("No", "Yes")),
      internet_phone = factor(as_factor(WP19544), levels = c("No", "Yes")),
      .keep = "unused",
      .before = 1
    )
  
  # Binary work
  dat$work <- as_factor(dat$work)
  count(dat, work)
  dat$work <- fct_collapse(
    dat$work, 
    no = c("Unemployed", "Out of workforce"), 
    yes = c(
      "Employed full time for an employer", 
      "Employed full time for self", 
      "Employed part time do not want full time", 
      "Employed part time want full time")
  )
  dat$work <- factor(dat$work, levels = c("no", "yes"), labels = c("No", "Yes"))
  
  # Fix one of each 13yr and 101yr
  dat$age <- if_else(dat$age < 15L, 15L, if_else(dat$age > 100L, 100L, dat$age))
  summary(dat$age)
  
  # primary index variables
  dat %>% 
    select(all_of(unlist(ix[1:4], use.names = FALSE))) %>% 
    generate_dictionary()
  dat <- dat %>% 
    mutate(
      # Life satisfaction is already in correct order but rescale
      life_satisfaction = if_else(
        WP16 %in% 98:99, 
        as.numeric(NaN), 
        as.numeric(WP16) / 10
      ),
      # Negative experiences, positive experiences, and "loneliness"
      # 1 is yes, 2 is no; reverse
      across(
        all_of(as.character(unlist(ix[2:4]))), 
        ~as.integer(2 - .)
      )
    ) %>% 
    mutate(
      negative_experiences = rowMeans(
        select(., WP68, WP69, WP70, WP71, WP74), na.rm = TRUE
      ),
      positive_experiences = rowMeans(
        select(., WP60, WP61, WP63, WP65, WP67), na.rm = TRUE
      ),
      social_life = rowMeans(
        select(., WP27, WP10248), na.rm = TRUE
      )
    )
  
  
  # Then the other scales
  dat[,-1] %>% 
    generate_dictionary()
  dat <- dat %>% 
    mutate(
      # Food and shelter [note below], 
      foodshelter = rowMeans(
        select(., WP40, WP43), na.rm = TRUE
      ) - 1,
      # GWPI items and scale to 0-1
      gwb_purpose = (
        rowMeans(select(., WP14442, WP14443), na.rm = TRUE) - 1
      ) / 4,
      gwb_community = (
        rowMeans(select(., WP14450, WP14451), na.rm = TRUE) - 1
      ) / 4,
      gwb_physical = (
        rowMeans(select(., WP14448, WP14449), na.rm = TRUE) - 1
      ) / 4,
      gwb_social = (
        rowMeans(select(., WP14444, WP14445), na.rm = TRUE) - 1
      ) / 4,
      .keep = "unused"
    )
  
  # Remove now unnecessary variables
  glimpse(dat)
  dat <- dat %>% 
    select(-starts_with("WP"))
  glimpse(dat)
  summary(dat)
  
  # A few dozen missing sex and age
  dat <- drop_na(dat, sex, age) %>% 
    mutate(sex = fct_drop(sex))
  
  # Income logp1
  dat <- mutate(dat, income = log(income+1))
  
  # Too little data in 2022, and poor internet value coverage in 2005
  dat <- filter(dat, between(year, 2006, 2021))
  
  # Create combined internet access variable
  dat <- dat %>% 
    mutate(
      internet_access = if_else(year < 2016, internet_home, internet_access)
    ) %>% 
    select(-internet_home)
  
  # Save
  write_rds(dat, DATA_PATH_GWP_CLEAN)
}
```

```{r}
#| label: data-synth-create

# Create synthetic smaller dataset for code reproducibility.
if (!file.exists(DATA_PATH_SYN)) {
  dir.create(dirname(DATA_PATH_SYN), FALSE)
  
  renv::settings$ignored.packages(c("synthpop", "mice"))
  library(synthpop)
  library(mice)
  
  set.seed(909090)
  data_synthetic <- read_rds(DATA_PATH_GWP_CLEAN) %>% 
    filter(country %in% sample(unique(country), 20)) %>% 
    slice_sample(n = 200, by = country) %>% 
    mice(m = 1) %>% 
    complete() %>% 
    tibble()
  
  data_synthetic <- data_synthetic %>% 
    syn() %>% 
    pluck("syn") %>% 
    tibble()

  write_rds(data_synthetic, DATA_PATH_SYN)
}
```


```{r}
#| label: data-load

if (file.exists(DATA_PATH_GWP_CLEAN)) {
  message("\n---\nUsing real data\n---\n")
  dat <- read_rds(DATA_PATH_GWP_CLEAN)
} else {
  message("\n---\nUsing synthetic data\n---\n")
  dat <- read_rds(DATA_PATH_SYN)
}

# Summarise original values
x_sum <- dat %>% 
  group_by(year) %>% 
  summarise(
    n = n(),
    countries = length(unique(country)),
    across(c(starts_with("internet")), ~sum(!is.na(.)))
  ) %>% 
  mutate(
    across(starts_with("internet"), ~ . / n)
  ) 

yr <- paste(range(dat$year), collapse = " to ")
```

```{r}
#| label: multiverse-specifications

# Select variables and categorize age
age_iqr <- quantile(dat$age, probs = c(.25, .75)) %>% paste(collapse = ", ")
dat <- dat %>% 
  select(
    country, year, 
    age, sex, 
    weight,
    education, work,
    healthproblems, relationship,
    foodshelter, income,
    starts_with("internet"),
    starts_with("gwb"),
    life_satisfaction, negative_experiences, 
    positive_experiences, social_life
  ) %>% 
  drop_na(age) %>% 
  mutate(
    age = cut(
      age, 
      breaks = c(15, 25, 35, 45, 55, 65, 101), 
      labels = c("15 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 +"),
      include.lowest = TRUE, right = FALSE
    )
  ) 

# Within-country center variables
dat_raw <- dat
dat <- dat %>% 
  group_by(country) %>% 
  mutate(
    across(
      c(education, income),
      ~. - mean(., na.rm = TRUE)
    ),
    across(
      c(healthproblems, foodshelter, relationship, work, starts_with("internet")), 
      ~((as.numeric(.)-1) - mean((as.numeric(.)-1), na.rm = TRUE))
    )
  ) %>% 
  ungroup()

#' Expand a vector of covariate names to all their combinations
#'
#' For example expand_covariate(c("age", "sex")) returns
#' c("age", "sex", "age + sex", "1")
#'
#' @param covariate e.g. c("age", "sex")
#'
#' @return a character vector
expand_covariate <- function(covariate) {
  list(
    "1",
    do.call(
      "c", 
      map(seq_along(covariate), ~combn(covariate, .x, FUN = list))) %>%
      map(~paste(.x, collapse = " + "))
  ) %>%
    unlist
}

No <- function(formula, data) {
  fit <- lmer(formula = formula, data = data, weights = rep(1, nrow(data)))
  out <- tidy(fit, conf.int = TRUE)
  bind_cols(out, n = nobs(fit))
}
Yes <- function(formula, data) {
  fit <- lmer(formula = formula, data = data, weights = weight)
  out <- tidy(fit, conf.int = TRUE)
  bind_cols(out, n = nobs(fit))
}
No <- possibly(No, tibble(n = 0))
Yes <- possibly(Yes, tibble(n = 0))

ys <- c(
    "life_satisfaction",
    "negative_experiences",
    "positive_experiences",
    "social_life",
    "gwb_purpose", 
    "gwb_physical",
    "gwb_community",
    "gwb_social"
  )
xs <- c(
    "internet_access",
    "internet_past7days",
    "internet_phone"
  )
cs <- c(
    "healthproblems",
    "relationship",
    "education",
    "work",
    "income",
    "foodshelter"
  ) %>% 
  expand_covariate()

specs <- expand_grid(
  distinct(dat, sex, age),
  y = ys,
  x = xs,
  covariate = cs,
  weights = c("No", "Yes")
) %>% 
  # Drop year from formula for GWB outcomes (only ~1 year)
  mutate(
    formula = if_else(
      str_starts(y, "gwb_"),
      str_glue("{y} ~ {x} + {covariate} + ({x} | country)"),
      str_glue("{y} ~ {x} + {covariate} + ({x} | country) + ({x} | year) + ({x} | country:year)"),
    )
  )

# If only simulated data exists, sun small subset of specs
set.seed(9091090)
if (DATA_PATH_GWP == "") specs <- slice_sample(specs, n = MAX_CORES)
```

```{r}
#| label: multiverse-run

path <- "tmp/multiverse-lmer.rds"
dir.create(dirname(path), FALSE)
if (!file.exists(path)) {
  
  library(furrr)
  plan(multisession, workers = MAX_CORES)
  
  # Pass these global objects to `future_pmap()`
  opts <- furrr_options(
    globals = list(
      dat = dat, No = No, Yes = Yes
    ),
    packages = c("lme4", "purrr", "dplyr", "broom.mixed")
  )
  
  specs <- specs %>% 
    mutate(
      out = future_pmap(
        list(weights, formula, sex, age),
        ~do.call(
          what = ..1, 
          args = list(
            formula = ..2, 
            data = filter(
              dat, 
              sex == ..3, 
              age == ..4
            )
          )
        ),
        .options = opts
      )
    ) %>% 
    unnest(out) %>% 
    filter(n > 0)
  
  saveRDS(specs, path)
  
}

specs <- readRDS(path)
```

```{r}
#| Label: multiverse-tidy

# Parameter of interest (slope of x)
specs <- specs %>% filter(term == x) 

# Arrange outcomes
specs <- specs %>% 
  mutate(y = factor(y, levels = ys))

# Arranged table with extra info 
mv <- specs %>% 
  arrange(estimate, y) %>% 
  group_by(y) %>% 
  mutate(i = 1:n()) %>% 
  ungroup() %>% 
  mutate(sig = sign(conf.low) == sign(conf.high)) %>% 
  mutate(
    decision = case_when(
      conf.low < 0 & conf.high < 0 ~ "Negative",
      conf.low > 0 & conf.high > 0 ~ "Positive",
      !sig ~ "Not significant"
    ) %>% 
      factor(levels = c("Negative", "Not significant", "Positive"))
  ) 

# Summary of multiverse magnitudes
mvs <- mv %>% 
  reframe(
    estimate = quantile(estimate, probs = c(.25, .5, .75)), 
    quantile = c(.25, .5, .75),
    .by = c(y)
  ) %>% 
  pivot_wider(
    names_from = quantile, 
    values_from = estimate
  )

# Summary of multiverse magnitudes and decisions
mvs <- mv %>% 
  mutate(
    decision = case_when(
      conf.low < 0 & conf.high < 0 ~ "Negative",
      conf.low > 0 & conf.high > 0 ~ "Positive",
      !sig ~ "Not significant"
    ) %>% 
      factor(levels = c("Negative", "Not significant", "Positive"))
  ) %>% 
  summarise(p = n(), .by = c(y, decision)) %>% 
  mutate(
    p = p / sum(p), .by = y
  ) %>% 
  complete(y, decision, fill = list(p = 0)) %>% 
  pivot_wider(names_from = decision, values_from = p) %>% 
  left_join(mvs) %>% 
  left_join(
    summarise(mv, n = median(n), .by = y)
  )

# Grand summary (also across outcome)
mvs_all <- mv %>% 
  mutate(
    estimate = if_else(y == "negative_experiences", -1*estimate, estimate)
  ) %>% 
  reframe(
    estimate = quantile(estimate, probs = c(.25, .5, .75)), 
    quantile = c(.25, .5, .75),
  ) %>% 
  pivot_wider(
    names_from = quantile, 
    values_from = estimate
  )

mvs_all <- mv %>% 
  mutate(
    across(
      c(estimate, conf.low, conf.high), 
      ~if_else(y == "negative_experiences", -1*., .)
    )
  ) %>% 
  mutate(
    decision = case_when(
      conf.low < 0 & conf.high < 0 ~ "Negative",
      conf.low > 0 & conf.high > 0 ~ "Positive",
      !sig ~ "Not significant"
    ) %>% 
      factor(levels = c("Negative", "Not significant", "Positive"))
  ) %>% 
  summarise(p = n(), .by = c(decision)) %>% 
  mutate(
    p = p / sum(p)
  ) %>% 
  complete(decision, fill = list(p = 0)) %>% 
  pivot_wider(names_from = decision, values_from = p) %>% 
  bind_cols(mvs_all) %>% 
  bind_cols(
    summarise(mv, n = median(n))
  )

mvs <- bind_rows(
  mvs,
  mutate(mvs_all, y = "Total")
) %>% 
  mutate(Outcome = sanitize_labels(y), N = n, .keep = "unused", .before = 1)
rm(mvs_all)

# Add count of specs
mvs <- mv %>% 
  count(Outcome = y, name = "K") %>% 
  add_row(Outcome = "Total", K = nrow(mv)) %>% 
  mutate(Outcome = sanitize_labels(Outcome)) %>% 
  left_join(mvs)
```

---
abstract: |
  Internet technologies' and platforms' potential psychological consequences remain debated. While these technologies have spurred new forms of commerce, education, and leisure, many are worried that they might negatively affect individuals by, for example, displacing time spent on other healthy activities. Relevant findings to date have been inconclusive and of limited geographic and demographic scope. We examined whether having (mobile) internet access or actively using the internet predicted eight well-being outcomes from `r yr` among `r number2(nrow(dat))` individuals across `r length(unique(dat$country))` countries. We first queried the extent to which well-being varied as a function of internet connectivity. Then, we examined these associations' robustness in a multiverse of `r number2(nrow(specs))` analysis specifications. `r percent(mvs$Positive[9], .1)` of these resulted in positive and statistically significant associations between internet connectivity and well-being. These results indicate that internet access and use predict well-being positively and independently from a set of plausible alternatives.
---

The increasing adoption and use of internet-enabled technologies and platforms has spurred debate about their potential effects on people's psychological well-being and functioning. Social scientists have shifted focus from other topics such as violent video games and television-based technologies to new and emerging platforms and handheld digital devices [@orbenSisypheanCycleTechnology2020]. Large technology firms such as Meta [@metaNewToolsResources2022], Google [@googleDigitalWellbeingTechnology2022], Apple [@appleIOS12Introduces2018], and TikTok [@tiktokDigitalWellbeing2021] have reacted to concerns and released a host of "digital well-being" tools. At the same time, health professionals [@officeofthesurgeongeneralosgProtectingYouthMental2021] and officials in many countries are working to enact new regulations [@departmentfordigitalculturemedia&sportGuideOnlineSafety2022] to ensure internet and technology platforms protect user well-being. However, even after considerable scientific attention, an understanding of the fundamental associations between internet technology adoption and use and well-being remains elusive, and results of scientific studies on this topic are decidedly mixed [@appelAreSocialMedia2020; @bestOnlineCommunicationSocial2014; @dickson2018screen].

The rise in tools, advice, and regulation aimed at addressing well-being is interesting, in part, because of how the studies informing this debate are done. For example, despite the fact the challenge is often framed as a world-wide issue, the geographic and demographic scope of the evidence base is narrow and not well mapped onto worldwide trends [@ghaiLackSampleDiversity2021]. In the past decades, the expansion of access to the internet has accelerated the most [@ituITUICTStatistics2021] in places where social scientists are the least likely to study their effects [@ghaiSocialMediaAdolescent2022]. Similarly, whilst studying technology adoption and well-being across countries requires frequent and high-quality measures of both factors, few if any investigations successfully combine the two. Measurement quality and consistency varies significantly and the most widely cited international studies either lack a longitudinal component [@byrneGlobalKidsOnline2016], or have long three [@oecdPISA2018Technical2018] or four year [@inchleyEnhancingUnderstandingAdolescent2020] intervals between data collections. Finally, because most of the debate surrounding the global impact of internet technologies is focused on younger people, little if any global data reflect associations between technology and well-being across the life course. This important lack of context means that it remains an open question of who, where, and when internet technologies and connectivity might be influencing people's well-being. Without knowing this it is impossible to deploy limited resources to capitalize on benefits or redress harms. To our knowledge, no research has directly grappled with these issues and addressed the worldwide scope of the debate.

Considering this impasse, our overarching research goal was to estimate the extent to which internet access, mobile internet access, and internet use might predict psychological well-being on a global level. To this end, we analyzed data from a series of cross-sectional samples of `r number2(nrow(dat))` individuals from `r length(unique(dat$country))` countries from `r yr`. We studied eight indicators of well-being; life satisfaction, the extent to which individuals reported experiencing daily negative and positive experiences; two indices of social well-being; physical well-being, community well-being, and experiences of purpose. We started by first asking how well-being differed between individuals who had access to or used (mobile) internet, and those who did not. To answer, we predicted the well-being indicators from whether individuals had access to the internet, smartphone access, or whether they were active internet users. We then sought to probe the robustness of these associations by conducting a series of multiverse analyses that examined how national, demographic, and individual factors might serve to mask or create links between internet connectivity and well-being.

# Methods

We analysed data from the Gallup World Poll [@gallupWorldPollMethodology2022]. Gallup World Poll (GWP) is a nationally representative (of each country's civilian, non-institutionalized, adult [15+] population) continuous annual panel survey of approximately 1,000 individuals from each of 164 countries, conducted from 2005 to 2022. Gallup conducts the surveys through one-hour interviews either face-to-face or via telephone, with questionnaires that are translated to the major conversational language of each country. See @gallupWorldPollMethodology2022 for the GWP methodological details. We used data from 2006 to 2021, because in 2005 the internet questions were answered by fewer than a quarter of respondents, and the 2022 data only contained five thousand responses from three countries. We refer to the `r yr` dataset below. The total sample size for this analysis was `r number2(nrow(dat))`, with `r length(unique(dat$country))` countries. The sample was `r percent(prop.table(table(dat$sex))["Female"], .1)` female, and the interquartile range of age was [`r age_iqr`].

## Internet access and use

The key variables that we considered as predictors measured the respondents' access and use of the internet. GWP has surveyed these with four items with varying coverage over time. First, GWP measured internet access with *"Does your home have access to the internet?"* from 2006 to 2015, where the mean percentage of non-missing values was `r percent(mean(x_sum$internet_home[1:10]), .1)`. From 2016 to 2021, a similar question measured overall internet access with *"Do you have access to the internet in any way, whether on a mobile phone, a computer, or some other device?"* (mean valid responses: `r percent(mean(x_sum$internet_access[11:16]), .1)`). Because these two questions were so similar, and to extend the range of data coverage, we merged these two to one variable indicating internet access.

Second, GWP asked about mobile internet access with *"Can your mobile phone be used to access the Internet?"* from 2017 to 2021 (mean valid responses: `r percent(mean(x_sum$internet_phone[12:16]), .1)`). Third, GWP measured internet use with *"Have you used the internet in the past seven days, whether on a mobile phone, a computer, or some other device?"* from 2015 to 2021 (mean valid responses: `r percent(mean(x_sum$internet_past7days[11:16]), .1)`). All of these items had binary "Yes"/"No" response options, and respondents were given the option to decline an answer or report that they didn't know. We coded the latter two as missing values.

## Well-being outcomes

We focused on eight measures of well-being, broadly defined, in the Gallup World Poll: Life satisfaction, negative and positive experiences, and social life satisfaction. In addition to these four outcomes, which were included in the GWP from 2006 to 2021, we studied four indicators from the Gallup-Sharecare Global Well-being Index™ (GWPI), which is a "barometer of individuals perceptions of their own well-being" [@gallupWorldPollMethodology2022].

Life satisfaction was measured with a single item: *"Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?"*, and respondents chose a number from 0 to 10. We rescaled this item to range from 0 to 1.

Negative and positive experiences are both measured through a set of five Yes/No items. For negative experiences, respondents answered the prompts *"Did you experience the following feelings during a lot of the day yesterday? How about physical pain / worry / sadness / stress / anger?"*. For positive experiences, the items were *"Did you feel well-rested yesterday?"*; *"Were you treated with respect all day yesterday?"*; *"Did you smile or laugh a lot yesterday?"*; *"Did you learn or do something interesting yesterday?"; and* "Did you experience the following feelings during a lot of the day yesterday? How about enjoyment?".

In addition, we included two variables intended to measure individuals' "social support structure and opportunities to make friends" (*"Someone in your life always encourages you to be healthy"* and *"Your friends and family give you positive energy every day"*). We included these items as a proxy for social well-being (hereafter, "social life"). While we recognize their potentially limited validity, we thought that, from the variables available, they best approximated an absence of loneliness, an important and much-studied aspect of well-being in connection to digital technologies.

In 2013 through 2015, the GWP database also included a (Gallup-Sharecare) Global Well-being Index™ (GWPI), which measured (among others) experiences of purpose (*"liking what one does each day and being motivated to achieve one's goals"*); community well-being (*"liking where one lives, feeling safe and having pride in one's community"*), physical well-being (*"having good health and enough energy to get things done daily"*), and social well-being (*"having supportive relationships and love in your life"*). Experiences of purpose were measured with prompts *"You like what you do every day"* and *"You learn or do something interesting every day"*. Community well-being was prompted with *"The city or area where you live is a perfect place for you"* and *"In the last 12 months, you have received recognition for helping to improve the city or area where you live"*. Physical well-being had prompts *"In the last seven days, you have felt active and productive every day"* and *"Your physical health is near-perfect"*. Social well-being was measured with the same items as the above "social life" index, but instead on a rating scale: All GWPI items included the prompt *"Thinking about your life in general, please rate your level of agreement with each of the following using a five-point scale, where 5 means you STRONGLY AGREE and 1 means you STRONGLY DISAGREE. You may choose any of the numbers 1, 2, 3, 4, or 5."*

While these scales are not psychometrically validated, Gallup cites prominent scientists as having helped with their development (@gallupWorldPollMethodology2022, p. 5). While this statement doesn't make up for these scales' lack of validation, we believe that the extensive scope of the dataset, across both time and countries, makes them uniquely valuable objects of study.

## Covariates

In addressing our second research question, our aim was to approach the independent contributions of internet access and use on well-being. To that end, we adjusted for plausible (and available) covariates in our models that might otherwise mask or bias any independent contributions of internet access on well-being, or create spurious associations. We chose six variables to represent such potentially confounding factors: The respondent's income, educational, work, and relationship statuses, their ability to meet basic needs for food and shelter, and whether or not they reported having health problems.

The GWP reports respondents' monthly household income in their local currencies. GWP used hot deck imputation for individuals who reported an income range (\~15%) or who did not provide responses (\~15%). GWP then converted those values to international dollars using the World Bank's purchasing power parity conversion factor, with the intent of making income estimates comparable across all respondents and countries. In our analyses, we log-transformed income. The GWP coded each respondent's educational status to one of three categories: Elementary education or less (up to eight years of education), some secondary education (nine to 15 years), and tertiary (education beyond high school). We used this variable as a continuous predictor. Since 2009, GWP measured work status with five levels related to the quantity of work, but we recoded this as a binary employed vs. not variable. Relationship status was measured with a question about marital status with six response options. We recoded this to a binary in relationship vs. not variable. Basic need satisfaction was measured with two items querying whether the respondent had had difficulties in providing for their food and shelter in the past 12 months. Respondents also answered "Do you have any health problems that prevent you from doing any of the things people your age normally can do?" with a binary yes/no response format.

In addition to these covariates, we identified meaningful subgroups in the data. Previous research has highlighted important differences between age groups and sexes in their levels of well-being and use of internet technologies, and associations between the two. We therefore conducted our analyses separately for each sex and age group. However, to reduce the computational complexity of our analyses, and because there are no strong a priori reasons to assume large differences between adjacent ages, we split the continuous age variable into six categories (15 to 24, 25 to 34, 35 to 44, 45 to 54, 55 to 64, and 65 +).

## Data analysis

Our general data-analytic approach was a regression model predicting one outcome (e.g. life satisfaction) from an intercept and one predictor (e.g. internet access). Where the outcome consisted of multiple items, we took the mean. In addition, we within-country centered all our predictors, so as to not include between-country differences in the coefficients, which consequently indicated contrasts between individuals who (e.g.) had access to the internet to those who didn't, within a given country. In addition, because the data were nested within countries, we specified a multilevel model where the intercept and coefficient of the internet predictor varied randomly over countries. To be conservative, we also added random intercepts and slopes over years, and the country by year interaction, because any associations might be heterogeneous over time [@vuorreThereNoEvidence2021]. We addressed our first research question about average contrasts with a multilevel model of well-being outcome $y$ for observation $i$, country $j$, year $k$ and year by country $l$, on internet predictor $x$, specified as

```{=tex}
\begin{align*} 
y_{ijkl} \sim\ &\text{Normal}(\mu_{ijkl}, \sigma^{2}), \\
\mu_{ijkl} =\ &\alpha_{0} + \beta_{0j} + \gamma_{0k} + \delta_{0l} + \\
  &(\alpha_{1} + \beta_{1j} + \gamma_{1k} + \delta_{1l})(x_{ijkl} - \bar{x}_{.jkl}), \\
\pmb{\beta} \sim\ &\text{Normal}\left(\begin{bmatrix}  0 \\ 0 \end{bmatrix}, \Sigma_\beta\right), \\
\pmb{\gamma} \sim\ &\text{Normal}\left(\begin{bmatrix}  0 \\ 0 \end{bmatrix}, \Sigma_\gamma\right), \\
\pmb{\delta} \sim\ &\text{Normal}\left(\begin{bmatrix}  0 \\ 0 \end{bmatrix}, \Sigma_\delta\right).
\end{align*}
```
In this manner, $\alpha_1$ is the difference in well-being indicator $y$ between individuals who had access to the internet and those who didn't, for the average country and year.

For our second, primary, research question we then conducted a multiverse analysis over different covariate permutations on this base model and subgroups in the data. We analysed all the possible ways in which the covariates could be included in this model (including no covariates) as fixed effects, leading to `r length(cs)` different covariate specifications. In addition to model specifications and subgroups, GWP recommends using model weights to adjust for demographic representativeness in some analyses, and so we conducted the analyses both with and without model weights. Our multiverse therefore consisted of all the distinct combinations of outcomes, predictors, age, sex, covariate combinations, and whether model weights were included or not, leading to `r number2(nrow(specs))` specifications. We used R [@rcoreteamLanguageEnvironmentStatistical2022] for analyses and the lme4 package for estimating the multilevel models [@batesFittingLinearMixedEffects2015].

# Data availability statement

We used the proprietary Gallup World Poll dataset, which is available to subscribing institutions through the Gallup website. All our code and a simulated GWP dataset is available at https://doi.org/10.5281/zenodo.7774923. We did not pre-specify the sample sizes, but instead used all the available data in the Gallup database (with the exclusions detailed below). This study was not pre-registered.

# Results

```{r}
#| label: multiverse-average-run

path <- "tmp/specs-lmer-total.rds"
if (!file.exists(path)) {
  
  library(furrr)
  plan(multisession, workers = 8)
  
  # Pass these global objects to `future_pmap()`
  opts <- furrr_options(
    globals = list(dat = dat, No = No),
    packages = c("lme4", "purrr", "dplyr", "broom.mixed")
  )
  
  specs_avg <- expand_grid(
    y = c(
      "life_satisfaction",
      "negative_experiences",
      "positive_experiences",
      "social_life",
      "gwb_purpose",
      "gwb_social",
      "gwb_physical",
      "gwb_community"
    ),
    x = c(
      "internet_access",
      "internet_past7days",
      "internet_phone"
    )
  ) %>% 
    # Drop year from formula for GWB outcomes (only ~1 year)
    mutate(
      model = "No", # No weights here
      formula = if_else(
        str_starts(y, "gwb_"),
        str_glue("{y} ~ {x} + ({x} | country)"),
        str_glue("{y} ~ {x} + ({x} | country) + ({x} | year) + ({x} | country:year)"),
      )
    ) %>% 
    mutate(
      out = future_pmap(
        list(model, formula),
        ~do.call(
          what = ..1, 
          args = list(
            formula = ..2, 
            data = dat
          )
        ),
        .options = opts
      )
    ) %>% 
    ungroup() %>% 
    unnest(out)

  specs_avg <- specs_avg %>% 
    filter(term == x) %>% 
    select(-c(effect, group, term, model, formula))
  
  saveRDS(specs_avg, path)
} else {specs_avg <- readRDS(path)}

# Median association (with reversed negative experiences)
# tmp <- specs_avg %>% 
#   mutate(estimate = if_else(y == "negative_experiences", -1*estimate, estimate)) %>% 
#   summarise(across(c(estimate, std.error), median))
```

```{r}
ysds <- summarise(dat, across(all_of(ys), ~sd(., na.rm = TRUE)))
stdz_internet_ls <- number(specs_avg$estimate[1] / ysds[1,1,TRUE], .01)
```

```{r}
#| label: fig-rq1
#| include: true
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.height: 5.6
#| fig.width: 9
#| fig.cap: Scatterplots of countries' mean well-being outcomes for individuals who responded No and Yes to the internet question, and their differences. Small points are countries' means, and solid colored points and lines are means of the countries' means. Numbers in each panel's left bottom corner indicate the regression coefficient (Yes - No) and its 95\% confidence interval. The GWBI outcomes did not overlap in time with mobile internet data.

tmp1 <- dat_raw %>% 
  pivot_longer(starts_with("internet")) %>% 
  drop_na(value) %>%
  group_by(country, name, value) %>% 
  summarise(
    across(
      all_of(ys), 
      ~mean(., na.rm = TRUE)
    ),
  ) %>% 
  ungroup() %>% 
  pivot_longer(contains("_"), names_to = "outcome", values_to = "y") %>% 
  filter(!(name == "internet_phone" & str_detect(outcome, "gwb_"))) %>%
  pivot_wider(names_from = value, values_from = y) %>% 
  mutate(z = Yes - No) %>% 
  pivot_longer(c(No, Yes, z), names_to = "x") %>% 
  mutate(outcome = factor(outcome, levels = ys))

tmp2 <- tmp1 %>% 
  group_by(name, outcome, x) %>% 
  summarise(m = mean_cl_boot(value, conf.int = 0.95)) %>% 
  ungroup() %>% 
  unnest(m) %>% 
  mutate(sig = sign(ymin) == sign(ymax)) %>% 
  mutate(
    decision = case_when(
      ymin < 0 & ymax < 0 ~ "Negative",
      ymin > 0 & ymax > 0 ~ "Positive",
      !sig ~ "Not significant"
    ) %>% 
      factor(levels = c("Negative", "Not significant", "Positive")) %>% 
      fct_drop()
  ) %>% 
  mutate(outcome = factor(outcome, levels = ys))

tmp1 %>% 
  ggplot(aes(x, value)) +
  scale_color_brewer(
    "Difference",
    palette = "Set1"
  ) +
  scale_y_continuous(
    "Well-being mean",
    breaks = extended_breaks(7),
    expand = expansion(.01)
  ) +
  scale_x_discrete(
    "Internet",
    expand = expansion(0.4),
    labels = ~str_replace(., "z", "Y-N")
  ) +
  geom_hline(yintercept = 0, linewidth = .1) +
  geom_line(
    data = . %>% filter(x != "z"),
    aes(group = country),
    linewidth = .1, alpha = .5, col = "gray70",
    position = position_quasirandom(width = .45)
  ) +
  geom_point(
    alpha = .55, size = .8, shape = 1, stroke = .33, 
    col = "gray70", fill = "white",
    position = position_quasirandom(width = .45)
  ) +
  geom_pointrange(
    data = tmp2 %>% filter(x %in% c("No", "Yes")),
    aes(y = y, ymin = ymin, ymax = ymax),
    size = 0.5, fatten = 1.8
  ) +
  geom_pointrange(
    data = tmp2 %>% filter(!(x %in% c("No", "Yes"))),
    aes(y = y, ymin = ymin, ymax = ymax, color = decision),
    size = 0.5, fatten = 1.5
  ) +
  geom_line(
    data = tmp2 %>% filter(x %in% c("No", "Yes")),
    aes(y = y),
    linewidth = .33,
    group = 1,
    show.legend = FALSE
  ) +
  geom_label(
    data = specs_avg %>% 
      mutate(
        across(c(where(is.numeric), -n), ~number(., .001)),
        n = number2(n),
        name = x, x = "No", outcome = y
      ) %>% 
      mutate(outcome = factor(outcome, levels = ys)),
    aes(
      label = str_glue("b = {estimate}\n[{conf.low}, {conf.high}]")
    ),
    x = -Inf, y = -Inf, 
    size = 2.5, hjust = 0, vjust = 0, lineheight = 1, fill = NA,
    label.size = 0, label.padding = unit(0.3, "line")
  ) +
  facet_grid(
    name ~ outcome, 
    labeller = labeller(
      .cols = ~sanitize_labels(.) %>% str_replace_all(" ", "\n"),
      .rows = ~sanitize_labels(.)
    )
  ) +
  theme(
    legend.position = "bottom"
  )
```

Our first research question concerned the average differences in well-being between individuals who had access to (mobile) internet or had used the internet in the past seven days, and those who didn't. We display the results of this analysis in Figure \@ref(fig:fig-rq1). The associations between internet access and well-being were consistently positive. For the average country, individuals who had access to the internet reported on average approximately 0.08 units greater life satisfaction, positive experiences, and social life satisfaction, and 0.06 units lower negative experiences than individuals who did not have access. Results regarding the more temporally restricted (2013-2015) GWBI outcomes portrayed a similar picture: Individuals with internet access reported approximately 0.08 units greater experiences of purpose, 0.1 unit greater physical, 0.02 units greater community, and 0.08 units greater social well-being than individuals without access.

Results across the other two internet access and use predictors were of the same sign and comparable magnitude. Being an active internet user was associated with 0.03 to 0.08 unit increases in life satisfaction, positive experiences, social well-being and physical well-being, and with a 0.04 unit decrease in negative experiences. The overall association between being an active internet user and community well-being was not significantly different from zero. Mobile phone internet access predicted increases between 0.06 and 0.07 units.

The estimates discussed above and in Figure \@ref(fig:fig-rq1) refer to percentages of the scale: To put these magnitudes to another context, the well-being outcomes' standard deviations across individuals, countries, and time ranged from `r number(min(ysds), .01)` to `r number(max(ysds), .01)`. In standardized terms, the observed differences were therefore quite small (e.g. the median life satisfaction difference was `r stdz_internet_ls` standard deviations between individuals who had access to the internet and those who did not), but not negligible.

Moreover, Figure \@ref(fig:fig-rq1) shows individual countries' observed difference scores as small points; they were very consistently in the same direction as the average contrasts, indicating that this difference held similarly across most countries. These results showed uniformly across the eight well-being outcomes that individuals who had access to, or actively used the internet reported meaningfully greater well-being than those who did not.

However, although the estimates in Figure \@ref(fig:fig-rq1) were surprisingly uniform across outcomes and predictors, they did not yet address our second, main research question: The robustness of this association across different analysis specifications and subgroups, and the extent to which internet access and use might independently predict individuals' well-being. To best answer this question, while at the same time appreciate the theoretical uncertainty in how, where, or for whom to study this association, we next turned to our main multiverse analysis.

```{r}
#| label: tbl-multiverse
#| include: true
mvs2 <- mvs %>% 
  mutate(
    K = number2(K, 1),
    N = number2(N, 1),
    across(c(Negative, `Not significant`, Positive), ~percent(., .1)),
    across(c(`0.25`, `0.5`, `0.75`), ~number(., .01))
  ) %>% 
  mutate(
    Median = `0.5`,
    IQR = str_glue("[{`0.25`}, {`0.75`}]"),
    .keep = "unused",
    .after = 3
  ) %>% 
  mutate(Outcome = factor(Outcome, levels = sanitize_labels(c(ys, "Total"))))
mvs2 %>%
  apa_table(
    caption = "Summary of multiverse analysis of associations between internet access (or use) and well-being.", 
    note = 'K is the number of specifications, and N is the median sample size across specifications. "Total" indicates quantities across well-being outcomes (we reversed associations with negative experiences for this aggregate number).', 
    col_spanners = list(
      "Association magnitude" = c(4, 5),
      "Proportions of significant associations" = c(6, 8)
    ), 
    font_size = "scriptsize",
    span_text_columns = TRUE
  )
```

```{r}
#| include: true
#| label: fig-multiverse
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.height: 6
#| fig.width: 9
#| fig.cap: Multiverse analysis results of associations between three internet use predictors and eight well-being outcomes. The estimated associations are ordered by increasing magnitude; the x-axis denotes the ordered number of (predictor * age * sex * covariate combination * weights) specification. The small shaded areas to the right of each panel show the kernel density estimates of the associations' point estimates, and the median association with a colored point. 
draw_sca <- function(mv, Y) {
  this_mv <- filter(mv, y %in% Y) %>% 
    mutate(y = factor(y, levels = Y))
  p1 <- this_mv %>% 
    ggplot(aes(i, estimate, col = decision)) +
    scale_color_manual(values = c("firebrick2", "grey70", "dodgerblue2")) +
    scale_y_continuous(
      position = "right",
      "Estimated association (95%CI)",
      breaks = extended_breaks(7)
    ) +
    scale_x_continuous(
      "Analysis specification (in order of association magnitude)",
      breaks = extended_breaks(5),
      expand = expansion(.015)
    ) +
    geom_hline(yintercept = 0, linewidth = .1) +
    geom_point(size = .25) +
    geom_linerange(
      aes(ymin = conf.low, ymax = conf.high),
      linewidth = .2, alpha = .2
    ) +
    geom_ysidedensity(
      aes(x = after_stat(ndensity)), 
      col = "grey20", fill = "grey60", alpha = .4, n = 3001
    ) +
    geom_ysidepoint(
      data = this_mv %>% 
        group_by(y) %>% 
        summarise(median = median(estimate)) %>% 
        # Just for coloring the dot
        mutate(
          decision = cut(
            median,
            breaks = c(-1, -.0001, .0001, 1),
            labels = c("Negative", "Not significant", "Positive")
          )
        ),
      aes(y = median, x = 0),
      size = 2, position = position_nudge(x = .025)
    ) +
    scale_ysidex_continuous(expand = expansion(c(0, .2))) +
    facet_wrap(
      "y", 
      scales = "free_x", 
      ncol = 4, 
      drop = TRUE,
      labeller = as_labeller(sanitize_labels)
    ) +
    theme(
      axis.title.x = element_blank(),
      legend.position = "none",
      ggside.panel.scale = 0.1,
      ggside.panel.spacing.x = unit(0, "pt"),
      ggside.axis.text = element_blank(),
      ggside.axis.ticks = element_blank()
    )
  
  p2 <- this_mv %>% 
    mutate(
      weights = factor(
        weights, 
        levels = c("No", "Yes"), 
        labels = c("No weights", "Weights"))
    ) %>% 
    pivot_longer(
      c(age, sex, x, covariate, weights), 
      values_transform = as.character
    ) %>%
    mutate(
      name = factor(
        name, 
        levels = c("x", "age", "sex", "covariate", "weights"),
        labels = c("X", "Age", " ", "Covariates", "  ")
      ),
      value = fct_relevel(value, cs),
      value = fct_recode(value, `No covariates` = "1")
    ) %>%
    ggplot(aes(i, value, col = decision)) +
    scale_color_manual(values = c("firebrick2", "grey70", "dodgerblue2")) +
    scale_y_discrete(
      position = "right",
      labels = ~sanitize_labels(., FALSE)
    ) +
    scale_x_continuous(
      "Analysis specification (ordered by increasing magnitude of association)",
      breaks = extended_breaks(5),
      expand = expansion(.01)
    ) +
    geom_point(
      shape = 20, 
      size = .1, 
      position = position_quasirandom(width = .2)
    ) +
    facet_grid(
      cols = vars(y),
      rows = vars(name), 
      scales = "free", 
      space = "free_y", 
      switch = "y", 
      drop = TRUE,
      labeller = as_labeller(sanitize_labels)
    ) +
    theme(
      axis.text.y = element_text(size = rel(0.6)),
      axis.title.y = element_blank(),
      legend.position = "none"
    )
  
  list(a = p1, b = p2)
}
tmp <- mv %>% 
  arrange(desc(std.error)) %>% 
  # Take out one degenerate estimate for this plot
  slice(-1) %>% 
  draw_sca(ys)
tmp$a +
  coord_cartesian(ylim = c(-0.1, 0.125)) +
  theme(axis.title.x = element_text())
```

We summarize the results of this multiverse analysis in Table \@ref(tab:tbl-multiverse). First, for life satisfaction, the median sample size across specifications was `r number2(mvs$N[[1]])` individuals. Answering 'Yes' to an internet access or use question was associated with a median `r number(mvs[["0.5"]][[1]], .01)` unit increase in life satisfaction. The central 50% of the distribution of associations was within the `r number(mvs[["0.25"]][[1]], .01)`, `r number(mvs[["0.75"]][[1]], .01)` interval. The association between internet access or use and life satisfaction was positive in `r percent(mvs$Positive[[1]], .01)` of model specifications. Numerical results for the other well-being outcomes were of similar magnitude, and the total proportion of specifications that resulted in negative associations was only `r percent(mvs$Negative[[9]], .01)`.

Figure \@ref(fig:fig-multiverse) shows that across all model specifications, the multilevel model's estimated association between internet access and use for the average country was very consistently positive. Each estimated association in Figure \@ref(fig:fig-multiverse) identifies a model specification (e.g. the coefficient for internet access for 15 to 24 year old males, adjusting for health problems and income, estimated with weights), which are ordered on increasing association magnitude.

Whereas Figure \@ref(fig:fig-rq1) showed that the overall average associations were consistent across internet adoption predictors and well-being outcomes, Figure \@ref(fig:fig-multiverse) highlights a surprising consistency in that association across demographic subgroups and, moreover, in model covariate specifications. If these associations were spurious associations caused by any of the covariates we considered, we should observe clusters of non-significant or reversed estimates for specific covariate sets. This was mostly not the case: Figures S1 and S2 show additional "dashboards" that indicate the covariate specifications for each estimate in Figure \@ref(fig:fig-multiverse), and we did not discern any clear patterns among them with respect to the covariate specifications.

```{r}
#| include: true
#| label: fig-multiverse-zoom
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.width: 9
#| fig.height: 5
#| fig.cap: Associations between the internet predictors and well-being outcomes by sex and age, ordered by covariate specification. The specifications are ordered such that greater numbers indicate more covariates in the model. Estimates are means across weighted and non-weighted models.
p1 <- specs %>% 
  mutate(y = factor(y, ys)) %>% 
  filter(term == x) %>% 
  summarise(
    estimate = mean(estimate),
    n = n(),
    .by = c(covariate, age, sex, x, y)
  ) %>%
  mutate(
    covariate = fct_relevel(covariate, unique(specs$covariate)),
    i = as.numeric(covariate)
  ) %>%
  ggplot(aes(i, estimate)) +
  geom_hline(yintercept = 0, size = .1) +
  scale_color_viridis_d(
    "Age"
  ) +
  scale_y_continuous(
    "Mean association"
  ) +
  scale_x_continuous(
    "Covariate specification"
  ) +
  guides(color = guide_legend(reverse = TRUE)) +
  geom_line(aes(group = age, col = age)) +
  facet_nested(
    x + sex ~ y,
    labeller = labeller(
      .cols = ~sanitize_labels(.) %>% str_replace_all(" ", "\n"),
      .rows = ~sanitize_labels(.)
    )
  )

p1
```

However, Figure \@ref(fig:fig-multiverse) shows a notable group of negative associations between community well-being and internet adoption among otherwise mostly positive relationships. To examine these differences further, Figure \@ref(fig:fig-multiverse-zoom) focuses on variability in the associations across demographics and covariate specifications. The association between community well-being and internet adoption was particular to active internet use (rather than access) and individuals in the youngest 15 to 24 age category.

More generally, Figure \@ref(fig:fig-multiverse-zoom) also shows that while increasing the number of covariates in the model did tend to decrease the magnitudes of the estimated associations, they typically remained positive. This observation shows that even after adjusting for all possible combinations of the covariates that we considered, the relationships between internet access or use and well-being remained positive. In turn, this suggests that the contributions of internet access and use on well-being were independent of the covariates we selected for, and thus might indicate causal relations. (Although we highlight the evidence here for causal claims is less than thin.)

# Discussion

The debate over internet platforms' and technologies' effects on individuals' psychological well-being remains a central topic because of their potentially global consequences. While past results on this topic have been mixed, the overwhelming majority of studies have examined convenience samples from the global north, thereby ignoring the fact that the penetration of the internet has been, and continues to be, a global phenomenon. In this study, we examined associations between internet use and access, and a broad variety of well-being indicators in a representative sample of `r number2(nrow(dat))` individuals across `r length(unique(dat$country))` countries within an age range that spanned from late adolescence to older adults. We found that on average across countries and demographics, individuals who had internet access, mobile internet access, or actively used the internet reported greater levels of life satisfaction, positive experiences, experiences of purpose, and physical, community, and social well-being, and lower levels of negative experiences.

The main thrust of our analysis was to examine to what extent this positive association between internet and well-being was sensitive to different demographic groups and modelling decisions. Furthermore, we attempted to estimate the unique contributions of internet on well-being by examining this association across `r length(cs)` different permutations of sets of plausible covariates that might otherwise create or mask associations. Across `r number2(nrow(specs))` of such model specifications, `r percent(mvs$Positive[9], .1)` of the associations were significantly positive, and only `r percent(mvs$Negative[9], .1)` were negative. These results therefore suggest that the proliferation of (mobile) internet access and increasing engagement with the internet has been a net benefit to the well-being of individuals around the world.

We did, however, observe a notable group of negative associations between internet use and community well-being. These negative associations were restricted to a portion of analyses concerning young (15-24 year old) individuals, and mostly females. These associations occurred across the full spectrum of covariate specifications, and thereby were not likely driven by a particular model specification. Further research should investigate whether low community well-being drives engagement with the internet, or vice versa.

Nevertheless, our conclusions are qualified by a number of factors. First, we compared individuals to each other. There are likely myriad other features about the human condition that are associated with both uptake of internet technologies and well-being in such a manner that they might cause positive associations to be observed. While we attempted to adjust for such features by including various covariates in our models, the data and theory to guide model selection were both limited.

Second, while between-person data such as we studied can inform inferences about average causal effects, longitudinal studies that track individuals and their internet use over time would be more informative in understanding the contexts of how and why an individual might be affected by internet technologies and platforms [@rohrerTheseAreNot2021].

Third, while the constructs that we studied represent the general gamut of well-being outcomes that are typically studied in connection to digital media and technology, they do not capture everything, nor are they standard and methodically validated measures otherwise found in the psychological literature. That is, the GWP data that we used represent a uniquely valuable resource in terms of its scope both over time and space. But the measurement quality of its items and scales might not be sufficient to capture the targeted constructs in the detailed manner that we would hope for. It is therefore possible that there are other features of well-being that are differently affected by internet technologies, and that our estimates might be noisier than would be found using psychometrically validated instruments.

Fourth, the validity of self-reported measures of technology engagement is found lacking, as self-reported quantities of use correlate only modestly with actual use, as measured for example by apps installed on smartphones [@parrySystematicReviewMetaanalysis2021; @wu-ouyangOverestimatingUnderestimatingCommunication2022]. In our study, we used reports of whether an individual has access to or has used the internet in the past week, which may suffer from these biases. Nevertheless, we believe it is more difficult to be mistaken in answering those questions than e.g. questions about average hours used in the past year.

Finally, we further highlight the tentative-at-most nature of our results with respect to causal effects of internet access and use on individuals' well-being. Causal inference from observational data, such as studied here, is notoriously difficult [@rohrerThinkingClearlyCorrelations2018; @vanderweeleCausalInferenceLongitudinal2016]. Critical theoretical assumptions must be made and properly applied in the statistical models in order to approach unbiased causal estimates, steps that we did not take in the current work. We nevertheless remain hopeful that the clarity with which we hoped to address this issue will provide a solid foundation for future work on internet technologies causal effects [@groszTabooExplicitCausal2020].

To overcome these limitations, we think the best way forward for this field is to expend more resources on collecting larger and more representative longitudinal datasets that include validated measurements of the constructs that researchers care about. In addition, these datasets should include accurate data on individuals' engagement with internet technologies in lieu of self-reports. Fortunately, both of these data are already collected; large cohort-based surveys in many countries already track individuals' psychological states over time, and internet platforms are infamous for collecting detailed data on their users' behaviors. A significant but potentially fruitful challenge then would be to marry those two streams of data and use them in transparent and independent scientific inquiry for a more detailed understanding of internet technologies in individuals' lives.

# Author contributions

Conceptualization: MV & AKP; Data curation: MV; Formal Analysis: MV; Funding acquisition: AKP; Investigation: MV & AKP; Methodology: MV & AKP; Project administration MV & AKP; Resources: AKP; Software: MV; Supervision: AKP; Visualization: MV & AKP; Writing -- original draft: MV & AKP; Writing -- review & editing: MV & AKP.

# References

::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\clearpage
\onecolumn
```
# Supplementary online material to *A multiverse analysis of the associations between internet use and well-being* (Vuorre & Przybylski, tbd)

<!-- Prepend table & figure numbers with S and reset counters -->

```{=tex}
\setcounter{page}{1}
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{SOM-\arabic{table}}
\renewcommand{\thefigure}{SOM-\arabic{figure}}
```
Table \@ref(tab:tbl-correlation-matrix) shows the pearson correlations between all variables, across demographics, time and space.

```{r}
#| label: tbl-correlation-matrix
#| include: true
x <- dat %>% 
  # filter(year > 2013) %>% 
  select(
    starts_with("internet"),
    income, healthproblems, relationship, foodshelter,
    life_satisfaction, negative_experiences, positive_experiences, social_life,
    starts_with("gwb_")
  ) %>% 
  mutate(across(where(is.factor), ~as.numeric(.)-1)) %>% 
  correlate()
y <- x %>% 
  # shave() %>% 
  fashion(na_print = " ") %>% 
  as_tibble() %>% 
  mutate(term = str_glue("{term} ({1:nrow(x)})")) %>% 
  mutate(term = sanitize_labels(term, FALSE))
names(y) <- c(" ", 1:(ncol(y)-1))
y %>%   
  apa_table(
    font_size = "scriptsize",
    placement = "h",
    caption = "Pearson correlations between key variables.",
    note = "We used all pairwise-complete observations."
  )
```

Figures \@ref(fig:fig-sca-full-1) and \@ref(fig:fig-sca-full-2) show the full specification curve analysis plots, including the "dashboards" in panels B.

```{r}
#| include: true
#| label: fig-sca-full-1
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.height: 9
#| fig.width: 9
#| fig.cap: Multiverse analysis results of associations between three internet use predictors and four main well-being outcomes. A. Estimated associations, in order of increasing magnitude, shown separately for each of the four main well-being outcomes. The small shaded areas to the right of each panel show the kernel density estimates of the associations' point estimates, and the median association with a colored point. B. A dashboard of model specifications. Each point in the small panels indicates the value of the grouping variable or covariate specification for the model indicated on the x-axis.
tmp <- draw_sca(mv, ys[1:4])
((tmp$a / tmp$b) +
      plot_layout(heights = c(3.2, 6.8))) +
    plot_annotation(tag_levels = "A")
```

```{r}
#| include: true
#| label: fig-sca-full-2
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.height: 9
#| fig.width: 9
#| fig.cap: Multiverse analysis results of associations between three internet use predictors and four GWBI well-being outcomes. A. Estimated associations, in order of increasing magnitude, shown separately for each of the four GWBI well-being outcomes. The small shaded areas to the right of each panel show the kernel density estimates of the associations' point estimates, and the median association with a colored point. B. A dashboard of model specifications. Each point in the small panels indicates the value of the grouping variable or covariate specification for the model indicated on the x-axis.
tmp <- draw_sca(mv, ys[5:8])
((tmp$a / tmp$b) +
      plot_layout(heights = c(3.2, 6.8))) +
    plot_annotation(tag_levels = "A")
```
